---
title: "Soil-Plant Water Balance and Groundwater - Laws, Bishop, Big Pine"
description: |
  On-off pumping management according to Green Book Section III values for Vegetation Water Requirements 1989-2026.
format: 
  html: 
    toc: false
    toc-depth: 3
    anchor-sections: true
    smooth-scroll: true
    code-fold: true
    code-summary: "code"
    code-line-numbers: true
    code-overflow: wrap
    code-link: true
    html-math-method: katex
tbl-cap-location: top   
# number-sections: true    
affiliation: "Inyo County Water Department"
date-modified: "2026-06-19"
citation:
  type: report
  container-title: "Green Book Section III On/Off Pumping Management"
  publisher: "Inyo County Water Department"
  issued: "2026-06-19"
  url: https://github.com/inyo-gov/pumping-management
google-scholar: true
---

```{r setup, include=FALSE}
library(lubridate)
library(dygraphs)
library(zoo)
library(readxl)
library(tidyverse)
library(here)

# source functions----
source(here('code','functions.R'))
source(here('code','input_schema.R'))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r read-data}
data_path <- here('data','Totals_Means_2025.csv')
tm2 <- read_csv(data_path) %>% filter(read != -777)

input_path <- here('data','onoff_input_05_2026.xlsx')
input <- read_onoff_input_workbook(input_path)
report_data <- as_report_inputs(input)

awc1 <- report_data$awc1
dtw1 <- report_data$dtw1
on.off1 <- report_data$on.off1
awc.req.turnon <- report_data$awc.req.turnon
linked_wells <- report_data$linked_wells
site_metadata <- report_data$site_metadata
```

```{r}
tozoo <- function(x) zoo(x$total_pumping, x$roy2)
dyMultiColumn <- function(dygraph) {
  dyPlotter(dygraph = dygraph,
            name = "MultiColumn",
            path = system.file("plotters/multicolumn.js",
                               package = "dygraphs"))
}

```

```{r}
plot_linked_wells <- function(data, site_id, linked_wells_df) {
  # Filter the linked_wells dataframe to only include wells linked to the specified site_id

    site_linked_wells <- linked_wells_df %>%
    filter(Site == site_id) %>% pull(Linked_Well)
  
  # Filter the data to include only the wells linked to the specified site_id
  # filtered_data <C- tm2 %>% 
    filtered_data <- data %>%
    filter(staid %in% site_linked_wells) %>%
    mutate(read2 = ifelse(read < 0, 0, read))%>%
    group_by(roy, staid) %>%
    summarise(total_pumping = sum(read)) %>%
    select(roy, staid, total_pumping) %>%
    mutate(roy2 = make_date(roy, 4, 1))
  
  # Replace NA values with 0
  filtered_data[is.na(filtered_data)] <- 0
  
  # Merge the filtered data by site_id
  merged_data <- do.call(merge, lapply(split(filtered_data, filtered_data$staid), tozoo))
  # Fill NA with 0 so bar heights match tooltip (avoids dygraphs interpolating NaN)
  .core <- coredata(merged_data)
  .core[is.na(.core)] <- 0
  coredata(merged_data) <- .core

  # Plot the merged data
  dygraph(merged_data, group = "a") %>%
    dyStackedBarGroup(site_linked_wells) %>%
    dyAxis("y", label = "Total Annual Pumping (AF)") %>%
    dyOptions(stackedGraph = TRUE)
}

# 
# print(filtered_data)
# merged_data
```

# On/Off Well Management

<details>
<summary>How on/off well management works</summary>

The Water Agreement and Green Book establish protocols for determining the operational status of LADWP pumping wells based on soil moisture and vegetation measurements. To fulfill monitoring requirements under the Agreement, the Water Department regularly assesses soil moisture content and vegetation leaf area index, in collaboration with LADWP, at 22 permanent sites within wellfields and seven sites in control areas. Soil moisture monitoring occurs monthly, while vegetation monitoring occurs annually in June. Each monitoring site is associated with one or more LADWP pumping wells.

On July 1 and October 1, if soil moisture levels are inadequate to support vegetation needs at a site, the linked pumping wells are placed in off status. They remain offline until soil moisture levels exceed the threshold required by the vegetation at the time the linked wells entered off status. This operational approach is known as the on/off management strategy.

Since the strategy began in 1990, LADWP has frequently maintained numerous wells in off status, consistent with subsequent agreements with the County.

</details>

# Soil-Plant Water Balance and Groundwater

<details>
<summary>How to read the site plots</summary>

Data for each permanent monitoring site (1990-present) are presented in three linked plots. The first plot shows pumping from linked production wells. The second shows available water content (AWC) and vegetation water requirements (VWR) over time. The third shows depth to water (DTW) from the associated monitoring well. Green shading marks periods when the site was in on status; off periods are unshaded.

Each site starts with a compact card showing the site photo, linked production wells, current on/off status, AWC required for well turn-on, and the associated monitoring well.

**Instructions**

-   `Zoom out`: Double-click a plot to return to the full time-series extent.
-   `Data point highlighting`: Move the cursor over a plot to display data values in the legend.
-   `Zoom in`: Drag horizontally to select a time period or vertically to select a custom y-axis range.
-   `Grouped selection`: Selecting a time period in one plot updates all plots.

</details>

```{r}

# Linked wells now come from the canonical input workbook.
invisible(linked_wells)

site_card <- function(
  site_id,
  title,
  image_src,
  image_alt,
  monitoring_well_label = NULL,
  map_src = NULL,
  map_alt = NULL
) {
  status_row <- awc.req.turnon[awc.req.turnon$site == site_id, , drop = FALSE]
  status_row <- status_row[1, , drop = FALSE]
  current_status <- status_row$current.status
  status_text <- ifelse(length(current_status) == 0 || is.na(current_status), "NA", current_status)
  status_class <- ifelse(status_text == "ON", "status-on", "status-off")
  turn_on_req <- status_row$AWC.req.turnon
  turn_on_text <- ifelse(length(turn_on_req) == 0 || is.na(turn_on_req), "NA", turn_on_req)
  wells_text <- linked_wells$Linked_Well[linked_wells$Site == site_id] %>%
    stringr::str_remove("^W") %>%
    paste(collapse = ", ")
  wells_text <- ifelse(nzchar(wells_text), wells_text, "Control site")

  htmltools::tags$section(
    class = ifelse(is.null(map_src), "site-card site-card-single", "site-card site-card-with-map"),
    htmltools::tags$div(
      class = ifelse(is.null(map_src), "site-card-media single-media", "site-card-media"),
      htmltools::tags$figure(
        class = "site-card-figure",
        htmltools::tags$img(
            class = "site-card-image site-card-photo-image",
          src = image_src,
          alt = image_alt
        ),
        htmltools::tags$figcaption("Site photo")
      ),
      if (!is.null(map_src)) {
        htmltools::tags$figure(
          class = "site-card-figure",
          htmltools::tags$img(
            class = "site-card-image site-card-map-image",
            src = map_src,
            alt = ifelse(is.null(map_alt), paste(title, "linked wells map"), map_alt)
          ),
          htmltools::tags$figcaption("Linked wells map")
        )
      }
    ),
    htmltools::tags$div(
      class = "site-card-body",
      htmltools::tags$p(class = "site-card-eyebrow", site_id),
      htmltools::tags$h4(class = "site-card-title", title),
      htmltools::tags$dl(
        class = "site-card-facts",
        htmltools::tags$div(
          htmltools::tags$dt("Current status"),
          htmltools::tags$dd(
            htmltools::tags$span(class = paste("site-status", status_class), status_text)
          )
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Linked pumping wells"),
          htmltools::tags$dd(wells_text)
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Soil AWC required for turn-on"),
          htmltools::tags$dd(turn_on_text)
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Monitoring well"),
          htmltools::tags$dd(ifelse(is.null(monitoring_well_label), "NA", monitoring_well_label))
        )
      )
    )
  )
}
```

### LAWS MONITORING SITE #1

```{r}
site_card(
  "LW1",
  "Laws Monitoring Site #1",
  "site_images/L1_0to100m.JPG",
  "Laws Monitoring Site #1 point-frame photo",
  monitoring_well_label = "795T",
  map_src = "site_images/site_maps/LW1_linked_wells_map.png",
  map_alt = "Static map showing Laws Monitoring Site #1 and linked wells W247, W248, W249, and W398"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "LW1",
  linked_wells,
  monitoring_well_label = "Depth to Water: 795T"
)
```

### LAWS MONITORING SITE #2

```{r}
site_card(
  "LW2",
  "Laws Monitoring Site #2",
  "site_images/LW2_0_2018.JPG",
  "Laws Monitoring Site #2 point-frame photo",
  monitoring_well_label = "V001G",
  map_src = "site_images/site_maps/LW2_linked_wells_map.png",
  map_alt = "Static map showing Laws Monitoring Site #2 and linked wells W236, W239, W243, and W244"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "LW2",
  linked_wells,
  monitoring_well_label = "Depth to Water: V001G"
)
```

### LAWS MONITORING SITE #3

```{r}
site_card(
  "LW3",
  "Laws Monitoring Site #3",
  "site_images/LW3_0_2018.JPG",
  "Laws Monitoring Site #3 point-frame photo",
  monitoring_well_label = "840T",
  map_src = "site_images/site_maps/LW3_linked_wells_map.png",
  map_alt = "Static map showing Laws Monitoring Site #3 and linked wells W240, W241, W399, W376, and W377"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "LW3",
  linked_wells,
  monitoring_well_label = "Depth to Water: 840T"
)
```

### BISHOP CONTROL SITE #1

```{r}
site_card(
  "BC1",
  "Bishop Control Site #1",
  "site_images/BC1_0to100m.JPG",
  "Bishop Control Site #1 point-frame photo",
  monitoring_well_label = "USGS 2AI"
)
```

```{r, fig.height=2.55}
plot_awc_dygraph(awc1,"BC1")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"BC1")

```

### BISHOP CONTROL SITE #2

```{r}
site_card(
  "BC2",
  "Bishop Control Site #2",
  "site_images/BC2_0_2018.JPG",
  "Bishop Control Site #2 point-frame photo",
  monitoring_well_label = "796T"
)
```

```{r, fig.height=2.55}
plot_awc_dygraph(awc1,"BC2")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"BC2")

```

### BISHOP CONTROL SITE #3

```{r}
site_card(
  "BC3",
  "Bishop Control Site #3",
  "site_images/BC3_0_2018.JPG",
  "Bishop Control Site #3 point-frame photo",
  monitoring_well_label = "797T"
)
```

```{r, fig.height=2.55}
plot_awc_dygraph(awc1,"BC3")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"BC3")

```

### BIG PINE MONITORING SITE #1

```{r}
site_card(
  "BP1",
  "Big Pine Monitoring Site #1",
  "site_images/BP1_0_2018.JPG",
  "Big Pine Monitoring Site #1 point-frame photo",
  monitoring_well_label = "798T",
  map_src = "site_images/site_maps/BP1_linked_wells_map.png",
  map_alt = "Static map showing Big Pine Monitoring Site #1 and linked wells W210, W378, W379, and W389"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "BP1",
  linked_wells,
  monitoring_well_label = "Depth to Water: 798T"
)
```

### BIG PINE MONITORING SITE #2

```{r}
site_card(
  "BP2",
  "Big Pine Monitoring Site #2",
  "site_images/BP2_0to100m.JPG",
  "Big Pine Monitoring Site #2 point-frame photo",
  monitoring_well_label = "799T",
  map_src = "site_images/site_maps/BP2_linked_wells_map.png",
  map_alt = "Static map showing Big Pine Monitoring Site #2 and linked wells W220, W229, W374, and W375"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "BP2",
  linked_wells,
  monitoring_well_label = "Depth to Water: 799T"
)
```

### BIG PINE MONITORING SITE #3

```{r}
site_card(
  "BP3",
  "Big Pine Monitoring Site #3",
  "site_images/BP3_0to100m.JPG",
  "Big Pine Monitoring Site #3 point-frame photo",
  monitoring_well_label = "567T",
  map_src = "site_images/site_maps/BP3_linked_wells_map.png",
  map_alt = "Static map showing Big Pine Monitoring Site #3 and linked wells W222, W223, W231, and W232"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "BP3",
  linked_wells,
  monitoring_well_label = "Depth to Water: 567T"
)
```

### BIG PINE MONITORING SITE #4

```{r}
site_card(
  "BP4",
  "Big Pine Monitoring Site #4",
  "site_images/BP4_0_2018.JPG",
  "Big Pine Monitoring Site #4 point-frame photo",
  monitoring_well_label = "800T",
  map_src = "site_images/site_maps/BP4_linked_wells_map.png",
  map_alt = "Static map showing Big Pine Monitoring Site #4 and linked well W331"
)
```

```{r}
plot_site_dashboard(
  tm2,
  awc1,
  dtw1,
  on.off1,
  "BP4",
  linked_wells,
  monitoring_well_label = "Depth to Water: 800T"
)
```
