---
title: "Soil-Plant Water Balance and Groundwater - Taboose-Aberdeen, Thibaut-Sawmill, Independence-Oak, Symmes-Shepherd, Bairs-Georges"
description: |
  On-off pumping management according to Green Book Section III values for Vegetation Water Requirements 1989-2026.
format: 
  html: 
    toc: false
    toc-depth: 3
    anchor-sections: true
    smooth-scroll: true
    code-fold: true
    code-summary: "code"
    code-line-numbers: true
    code-overflow: wrap
    code-link: true
    html-math-method: katex
tbl-cap-location: top   
# number-sections: true    
affiliation: "Inyo County Water Department"
date-modified: "2026-06-19"
citation:
  type: report
  container-title: "Green Book Section III On/Off Pumping Management"
  publisher: "Inyo County Water Department"
  issued: "2026-06-19"
  url: https://github.com/inyo-gov/pumping-management
google-scholar: true
---

```{r setup, include=FALSE}
library(lubridate)
library(dygraphs)
library(zoo)
library(readxl)
library(tidyverse)
library(here)

# source functions----
source(here('code','functions.R'))
source(here('code','input_schema.R'))

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r read-data}
data_path <- here('data','Totals_Means_2025.csv')
tm2 <- read_csv(data_path) %>% filter(read != -777)

input_path <- here('data','onoff_input_05_2026.xlsx')
input <- read_onoff_input_workbook(input_path)
report_data <- as_report_inputs(input)

awc1 <- report_data$awc1
dtw1 <- report_data$dtw1
on.off1 <- report_data$on.off1
awc.req.turnon <- report_data$awc.req.turnon
linked_wells <- report_data$linked_wells
site_metadata <- report_data$site_metadata
```

# On/Off Well Management

<details>
<summary>How on/off well management works</summary>

The Water Agreement and Green Book establish protocols for determining the operational status of LADWP pumping wells based on soil moisture and vegetation measurements. To fulfill monitoring requirements under the Agreement, the Water Department regularly assesses soil moisture content and vegetation leaf area index, in collaboration with LADWP, at 22 permanent sites within wellfields and seven sites in control areas. Soil moisture monitoring occurs monthly, while vegetation monitoring occurs annually in June. Each monitoring site is associated with one or more LADWP pumping wells.

On July 1 and October 1, if soil moisture levels are inadequate to support vegetation needs at a site, the linked pumping wells are placed in off status. They remain offline until soil moisture levels exceed the threshold required by the vegetation at the time the linked wells entered off status. This operational approach is known as the on/off management strategy.

Since the strategy began in 1990, LADWP has frequently maintained numerous wells in off status, consistent with subsequent agreements with the County.

</details>

# Soil-Plant Water Balance and Groundwater

<details>
<summary>How to read the site plots</summary>

Data for each permanent monitoring site (1990-present) are presented in three linked plots. The first plot shows pumping from linked production wells. The second shows available water content (AWC) and vegetation water requirements (VWR) over time. The third shows depth to water (DTW) from the associated monitoring well. Green shading marks periods when the site was in on status; off periods are unshaded.

Each site starts with a compact card showing the site photo, linked production wells, current on/off status, AWC required for well turn-on, the associated monitoring well, and a static map of the site and linked wells.

**Instructions**

-   `Zoom out`: Double-click a plot to return to the full time-series extent.
-   `Data point highlighting`: Move the cursor over a plot to display data values in the legend.
-   `Zoom in`: Drag horizontally to select a time period or vertically to select a custom y-axis range.
-   `Grouped selection`: Selecting a time period in one plot updates all plots.

</details>

```{r}
# Linked wells now come from the canonical input workbook.
invisible(linked_wells)

site_card <- function(
  site_id,
  title,
  image_src,
  image_alt,
  monitoring_well_label = NULL,
  map_src = NULL,
  map_alt = NULL
) {
  status_row <- awc.req.turnon[awc.req.turnon$site == site_id, , drop = FALSE]
  status_row <- status_row[1, , drop = FALSE]
  current_status <- status_row$current.status
  status_text <- ifelse(length(current_status) == 0 || is.na(current_status), "NA", current_status)
  status_class <- ifelse(status_text == "ON", "status-on", "status-off")
  turn_on_req <- status_row$AWC.req.turnon
  turn_on_text <- ifelse(length(turn_on_req) == 0 || is.na(turn_on_req), "NA", turn_on_req)
  wells_text <- linked_wells$Linked_Well[linked_wells$Site == site_id] %>%
    stringr::str_remove("^W") %>%
    paste(collapse = ", ")
  wells_text <- ifelse(nzchar(wells_text), wells_text, "Control site")

  htmltools::tags$section(
    class = ifelse(is.null(map_src), "site-card site-card-single", "site-card site-card-with-map"),
    htmltools::tags$div(
      class = ifelse(is.null(map_src), "site-card-media single-media", "site-card-media"),
      htmltools::tags$figure(
        class = "site-card-figure",
        htmltools::tags$img(
            class = "site-card-image site-card-photo-image",
          src = image_src,
          alt = image_alt
        ),
        htmltools::tags$figcaption("Site photo")
      ),
      if (!is.null(map_src)) {
        htmltools::tags$figure(
          class = "site-card-figure",
          htmltools::tags$img(
            class = "site-card-image site-card-map-image",
            src = map_src,
            alt = ifelse(is.null(map_alt), paste(title, "linked wells map"), map_alt)
          ),
          htmltools::tags$figcaption("Linked wells map")
        )
      }
    ),
    htmltools::tags$div(
      class = "site-card-body",
      htmltools::tags$p(class = "site-card-eyebrow", site_id),
      htmltools::tags$h4(class = "site-card-title", title),
      htmltools::tags$dl(
        class = "site-card-facts",
        htmltools::tags$div(
          htmltools::tags$dt("Current status"),
          htmltools::tags$dd(
            htmltools::tags$span(class = paste("site-status", status_class), status_text)
          )
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Linked pumping wells"),
          htmltools::tags$dd(wells_text)
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Soil AWC required for turn-on"),
          htmltools::tags$dd(turn_on_text)
        ),
        htmltools::tags$div(
          htmltools::tags$dt("Monitoring well"),
          htmltools::tags$dd(ifelse(is.null(monitoring_well_label), "NA", monitoring_well_label))
        )
      )
    )
  )
}
```

### TABOOSE/ABERDEEN MONITORING SITE #3

```{r}
site_card("TA3", "Taboose/Aberdeen Monitoring Site #3", "site_images/TA3_0to100m.JPG", "Taboose/Aberdeen Monitoring Site #3 point-frame photo", monitoring_well_label = "849T", map_src = "site_images/site_maps/TA3_linked_wells_map.png", map_alt = "Static map showing Taboose/Aberdeen Monitoring Site #3 and linked wells W106, W110, W111, and W114")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TA3", linked_wells, monitoring_well_label = "Depth to Water: 849T")
```

### TABOOSE/ABERDEEN MONITORING SITE #4

```{r}
site_card("TA4", "Taboose/Aberdeen Monitoring Site #4", "site_images/TA4_0_2018.JPG", "Taboose/Aberdeen Monitoring Site #4 point-frame photo", monitoring_well_label = "586T", map_src = "site_images/site_maps/TA4_linked_wells_map.png", map_alt = "Static map showing Taboose/Aberdeen Monitoring Site #4 and linked wells W342 and W347")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TA4", linked_wells, monitoring_well_label = "Depth to Water: 586T")
```

### TABOOSE/ABERDEEN MONITORING SITE #5

```{r}
site_card("TA5", "Taboose/Aberdeen Monitoring Site #5", "site_images/TA5_0_2018.JPG", "Taboose/Aberdeen Monitoring Site #5 point-frame photo", monitoring_well_label = "801T", map_src = "site_images/site_maps/TA5_linked_wells_map.png", map_alt = "Static map showing Taboose/Aberdeen Monitoring Site #5 and linked well W349")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TA5", linked_wells, monitoring_well_label = "Depth to Water: 801T")
```

### TABOOSE/ABERDEEN MONITORING SITE #6

```{r}
site_card("TA6", "Taboose/Aberdeen Monitoring Site #6", "site_images/TA6_0_2018.JPG", "Taboose/Aberdeen Monitoring Site #6 point-frame photo", monitoring_well_label = "803T", map_src = "site_images/site_maps/TA6_linked_wells_map.png", map_alt = "Static map showing Taboose/Aberdeen Monitoring Site #6 and linked wells W109 and W370")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TA6", linked_wells, monitoring_well_label = "Depth to Water: 803T")
```

### THIBAUT/SAWMILL MONITORING SITE #1

```{r}
site_card("TS1", "Thibaut/Sawmill Monitoring Site #1", "site_images/TS1_0to100m.JPG", "Thibaut/Sawmill Monitoring Site #1 point-frame photo", monitoring_well_label = "807T", map_src = "site_images/site_maps/TS1_linked_wells_map.png", map_alt = "Static map showing Thibaut/Sawmill Monitoring Site #1 and linked well W159")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TS1", linked_wells, monitoring_well_label = "Depth to Water: 807T")
```

### THIBAUT/SAWMILL MONITORING SITE #2

```{r}
site_card("TS2", "Thibaut/Sawmill Monitoring Site #2", "site_images/TS2_0_2018.JPG", "Thibaut/Sawmill Monitoring Site #2 point-frame photo", monitoring_well_label = "806T", map_src = "site_images/site_maps/TS2_linked_wells_map.png", map_alt = "Static map showing Thibaut/Sawmill Monitoring Site #2 and linked well W155")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TS2", linked_wells, monitoring_well_label = "Depth to Water: 806T")
```

### THIBAUT/SAWMILL MONITORING SITE #3

```{r}
site_card("TS3", "Thibaut/Sawmill Monitoring Site #3", "site_images/TS3_0to100m.JPG", "Thibaut/Sawmill Monitoring Site #3 point-frame photo", monitoring_well_label = "851T", map_src = "site_images/site_maps/TS3_linked_wells_map.png", map_alt = "Static map showing Thibaut/Sawmill Monitoring Site #3 and linked wells W103, W104, and W382")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TS3", linked_wells, monitoring_well_label = "Depth to Water: 851T")
```

### THIBAUT/SAWMILL MONITORING SITE #4

```{r}
site_card("TS4", "Thibaut/Sawmill Monitoring Site #4", "site_images/TS4_0to100m.JPG", "Thibaut/Sawmill Monitoring Site #4 point-frame photo", monitoring_well_label = "804T", map_src = "site_images/site_maps/TS4_linked_wells_map.png", map_alt = "Static map showing Thibaut/Sawmill Monitoring Site #4 and linked wells W380 and W381")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "TS4", linked_wells, monitoring_well_label = "Depth to Water: 804T")
```

### THIBAUT/SAWMILL CONTROL SITE

```{r}
site_card("TSC", "Thibaut/Sawmill Control Site", "site_images/TSC_0_2018.JPG", "Thibaut/Sawmill Control Site point-frame photo", monitoring_well_label = "805T")
```

```{r, fig.height=2.55}

# AWC and VWR
plot_awc_dygraph(awc1,"TSC")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"TSC")
 
```

### INDEPENDENCE/OAK MONITORING SITE #1

```{r}
site_card("IO1", "Independence/Oak Monitoring Site #1", "site_images/IO1_0to100m.JPG", "Independence/Oak Monitoring Site #1 point-frame photo", monitoring_well_label = "809T", map_src = "site_images/site_maps/IO1_linked_wells_map.png", map_alt = "Static map showing Independence/Oak Monitoring Site #1 and linked wells W061, W391, and W400")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "IO1", linked_wells, monitoring_well_label = "Depth to Water: 809T")
```

### INDEPENDENCE/OAK MONITORING SITE #2

```{r}
site_card("IO2", "Independence/Oak Monitoring Site #2", "site_images/IO2_0_2018.JPG", "Independence/Oak Monitoring Site #2 point-frame photo", monitoring_well_label = "854T", map_src = "site_images/site_maps/IO2_linked_wells_map.png", map_alt = "Static map showing Independence/Oak Monitoring Site #2 and linked well W063")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "IO2", linked_wells, monitoring_well_label = "Depth to Water: 854T")
```

### INDEPENDENCE/OAK CONTROL SITE #1

```{r}
site_card("IC1", "Independence/Oak Control Site #1", "site_images/IC1_0to100m.JPG", "Independence/Oak Control Site #1 point-frame photo", monitoring_well_label = "USGS 8-D")
```

```{r, fig.height=2.55}

# AWC and VWR
plot_awc_dygraph(awc1,"IC1")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"IC1")

```

### INDEPENDENCE/OAK CONTROL SITE #2

```{r}
site_card("IC2", "Independence/Oak Control Site #2", "site_images/IC2_0_2018.JPG", "Independence/Oak Control Site #2 point-frame photo", monitoring_well_label = "810T")
```

```{r, fig.height=2.55}

# AWC and VWR
plot_awc_dygraph(awc1,"IC2")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"IC2")

```

### SYMMES/SHEPHERD MONITORING SITE #1

```{r}
site_card("SS1", "Symmes/Shepherd Monitoring Site #1", "site_images/SS1_0to100m.JPG", "Symmes/Shepherd Monitoring Site #1 point-frame photo", monitoring_well_label = "V009G", map_src = "site_images/site_maps/SS1_linked_wells_map.png", map_alt = "Static map showing Symmes/Shepherd Monitoring Site #1 and linked wells W069, W392, and W393")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "SS1", linked_wells, monitoring_well_label = "Depth to Water: V009G")
```

### SYMMES/SHEPHERD MONITORING SITE #2

```{r}
site_card("SS2", "Symmes/Shepherd Monitoring Site #2", "site_images/SS2_0_2018.JPG", "Symmes/Shepherd Monitoring Site #2 point-frame photo", monitoring_well_label = "646T", map_src = "site_images/site_maps/SS2_linked_wells_map.png", map_alt = "Static map showing Symmes/Shepherd Monitoring Site #2 and linked wells W074, W394, and W395")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "SS2", linked_wells, monitoring_well_label = "Depth to Water: 646T")
```

### SYMMES/SHEPHERD MONITORING SITE #3

```{r}
site_card("SS3", "Symmes/Shepherd Monitoring Site #3", "site_images/SS3_0to100m.JPG", "Symmes/Shepherd Monitoring Site #3 point-frame photo", monitoring_well_label = "561T", map_src = "site_images/site_maps/SS3_linked_wells_map.png", map_alt = "Static map showing Symmes/Shepherd Monitoring Site #3 and linked wells W092 and W396")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "SS3", linked_wells, monitoring_well_label = "Depth to Water: 561T")
```

### SYMMES/SHEPHERD MONITORING SITE #4

```{r}
site_card("SS4", "Symmes/Shepherd Monitoring Site #4", "site_images/SS4_0_2018.JPG", "Symmes/Shepherd Monitoring Site #4 point-frame photo", monitoring_well_label = "885T", map_src = "site_images/site_maps/SS4_linked_wells_map.png", map_alt = "Static map showing Symmes/Shepherd Monitoring Site #4 and linked wells W075 and W345")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "SS4", linked_wells, monitoring_well_label = "Depth to Water: 885T")
```

### BAIRS GEORGES MONITORING SITE #2

```{r}
site_card("BG2", "Bairs/Georges Monitoring Site #2", "site_images/BG2_0to100m.JPG", "Bairs/Georges Monitoring Site #2 point-frame photo", monitoring_well_label = "812T", map_src = "site_images/site_maps/BG2_linked_wells_map.png", map_alt = "Static map showing Bairs/Georges Monitoring Site #2 and linked wells W076, W403, W343, and W348")
```

```{r}
plot_site_dashboard(tm2, awc1, dtw1, on.off1, "BG2", linked_wells, monitoring_well_label = "Depth to Water: 812T")
```

### BAIRS/GEORGES CONTROL SITE

```{r}
site_card("BGC", "Bairs/Georges Control Site", "site_images/BGC_0to100m.JPG", "Bairs/Georges Control Site point-frame photo", monitoring_well_label = "856T")
```

```{r, fig.height=2.55}

# AWC and VWR
plot_awc_dygraph(awc1,"BGC")

```

```{r, fig.height=2.55}
# DTW
plot_dtw_dygraph(dtw1,"BGC")

```

